---
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: false
    theme: paper
    code_folding: none
    toc_float:
        collapsed: false
        smooth_scroll: false
---

```{r setup, echo=FALSE}
suppressPackageStartupMessages({
  suppressWarnings({
    library(dplyr)
    library(tibble)
    library(tidyr)
    library(readr)
    library(stringr)
    library(here)
    library(scales)
    library(log4r)
    library(lubridate)
    library(purrr)
    library(arrow)
    library(tidyjson)
    library(ggplot2)
    library(ggrepel)
    library(patchwork)
  })
})

mainfont <- "Source Sans Pro"
tab <- function (x = c("blue", "orange", "red", "seablue", "green", 
                       "olive", "purple", "pink", "brown", "gray")) {
  tableau <- c(blue = "#4E79A7", orange = "#F28E2B", red = "#E15759", seablue = "#76B7B2", 
               green = "#59A14F", olive = "#EDC948", purple = "#B07AA1", pink = "#FF9DA7", 
               brown = "#9C755F", gray = "#BAB0AC")
  as.vector(tableau[x])
}

theme_set(theme_minimal(base_family=mainfont) +
            theme(panel.grid = element_line(color="gray95"),
                  plot.caption = element_text(size=6, face = "italic", color="gray60"),
                  text = element_text(family=mainfont),
                  legend.title = element_blank(),
                  legend.position="bottom"))
update_geom_defaults('col', list(fill=tab('blue')))
update_geom_defaults('bar', list(fill=tab('blue')))
update_geom_defaults('text', list(family=mainfont, size=8/.pt))
update_geom_defaults('label', list(family=mainfont, size=8/.pt, label.size=NA))
# update_geom_defaults('point', list(shape=21, size=2, fill=tab('blue'), color='white'))
update_geom_defaults('point', list(size=0.5, color=tab('blue')))
update_geom_defaults('line', list(color=tab('blue'), size=0.85))

knitr::opts_chunk$set(
  cache = FALSE,
  message = FALSE,
  warning = FALSE, 
  dev = c("png", "cairo_pdf"),
  echo = FALSE,
  fig.retina = 2,
  fig.width = 7,
  fig.height = 3.5
)
img_slug <- "cna_metrics"
```

# containers.cna.metrics 

```{r}

mklab <- function (x, bytes = FALSE) {
    brate <- tibble(cut = c(10^seq(3, 24, 3), Inf), suffix = c("0", 
        "K", "M", "B", "T", "q", "Q", "s", "S"))
    if (bytes) {
        brate <- tibble(cut = c(1024^seq(8), Inf), suffix = c("0", 
            "kB", "MB", "GB", "TB", "PB", "EB", "ZB", "YB"))
    }
    pickup <- tibble(x = x) %>% mutate(one = cut(x, breaks = c(-Inf, 
        brate$cut), labels = brate$suffix, right = FALSE)) %>% 
        mutate(two = x/sapply(.data$one, function(y) lag(brate$cut)[y])) %>% 
        mutate(three = ifelse(.data$two < 10, as.character(round(.data$two, 
            1)), as.character(round(.data$two, 0)))) %>% mutate(three = ifelse(is.na(.data$three), 
        as.character(round(.data$x, 0)), paste0(.data$three, 
            .data$one))) %>% mutate(four = ifelse(x < 1, x, .data$three))
    pickup$four
}

```

## Count of CVEs and CNAs use of containers.cna.metrics

```{r}
cves <- read_parquet(here("cache/cveparse/cves.parquet")) %>% 
  select(-contains(".x_"))
cvepub <- cves %>% 
  select(cve, published = cveMetadata.datePublished) %>% drop_na() %>% 
  mutate(published = as.Date(str_extract(published, "\\d{4}-\\d{2}-\\d{2}")))
cvemonth <- cvepub %>% 
  mutate(month = floor_date(published, unit="month")) %>% 
  count(month)
cvecna <- cves %>% 
  select(cve, cna = cveMetadata.assignerShortName) %>% 
  drop_na()
cnacount <- cves %>% 
  select(cve, cna = cveMetadata.assignerShortName) %>% 
  drop_na() %>%
  count(cna, name="n")
totalcve <- cves %>% distinct(cve) %>% nrow()
totalcna <- cves %>% distinct(cveMetadata.assignerShortName) %>% nrow()


raw_metrics <- read_parquet(here("cache", "cveparse", "containers.cna.metrics.parquet")) %>% 
  left_join(read_parquet(here("cache", "cveparse", "containers.cna.metrics.scenarios.parquet")), by=c("cve", "metrics_id"))

metrics <- raw_metrics %>% 
  select(cve, metrics_id, ends_with(c("vectorString", "version", "baseScore"))) %>% 
  distinct() %>% 
  pivot_longer(cols = -c(cve, metrics_id), names_to="key", values_to = "val", 
               values_transform = as.character, values_drop_na = TRUE) %>% 
  mutate(key = gsub("containers.cna.metrics.", "", key)) %>% 
  separate(key, into=c("key_version", "key_name"), sep="\\.") %>% 
  mutate(key_version = case_when(key_version == "cvssV3_1" ~ "3.1",
                                 key_version == "cvssV3_0" ~ "3.0",
                                 key_version == "cvssV2_0" ~ "2.0",
                                 TRUE ~ "unknown")) %>% 
  pivot_wider(id_cols = c(cve, metrics_id, key_version), names_from = key_name, values_from = val)


metrics_count <- raw_metrics %>% 
  pivot_longer(cols = -c(cve, metrics_id, scenarios_id), names_to="key", values_to = "val", 
               values_transform = as.character, values_drop_na = TRUE) %>% 
  left_join(cvecna, by=c("cve")) %>% 
  mutate(key = str_replace(key, "containers.cna.", "")) %>% 
  summarize(.by=key, cves = n_distinct(cve), cnas = n_distinct(cna)) %>% 
  mutate(dots = str_count(key, "\\.")) %>% 
  arrange(dots, key) %>% 
  select(-dots)

knitr::kable(metrics_count, format = "html", 
             caption = paste(cves %>% distinct(cve) %>% nrow() %>% comma, 
                     "CVEs from", comma(totalcna), "CNAs as of ", 
                     as.Date(max(cves$cveMetadata.datePublished, na.rm = TRUE))))

```


## Percent of Published CVEs with CVSS information by month

Each dot in the plot below is a month and is showing the percent of CVEs published in that month that had at least one complete combination of a CVSS vector string and base score published in the CVE record. 

```{r}
metrics <- raw_metrics %>% 
  select(cve, metrics_id, ends_with(c("vectorString", "version", "baseScore"))) %>% 
  distinct() %>% 
  pivot_longer(cols = -c(cve, metrics_id), names_to="key", values_to = "val", 
               values_transform = as.character, values_drop_na = TRUE) %>% 
  mutate(key = gsub("containers.cna.metrics.", "", key)) %>% 
  separate(key, into=c("key_version", "key_name"), sep="\\.") %>% 
  mutate(key_version = case_when(key_version == "cvssV3_1" ~ "3.1",
                                 key_version == "cvssV3_0" ~ "3.0",
                                 key_version == "cvssV2_0" ~ "2.0",
                                 TRUE ~ "unknown")) %>% 
  pivot_wider(id_cols = c(cve, metrics_id, key_version), names_from = key_name, values_from = val) %>% 
  full_join(cvepub, by="cve") %>% 
  full_join(cvecna, by="cve") %>%
  mutate(has_cvss = !is.na(vectorString)) %>% 
  summarize(.by=c(cve, published, cna), has_cvss = ifelse(any(has_cvss), "yes", "no"))
  # filter(is.na(vectorString) | is.na(version) | is.na(baseScore) | is.na(key_version) | is.na(published) | is.na(cna))

toplot <- metrics %>% 
  mutate(month = floor_date(published, unit="month")) %>% 
  count(month, has_cvss) %>% 
  spread(has_cvss, n, fill=0) %>% 
  mutate(total = no + yes, pct = yes / total) %>%
  mutate(colfill = ifelse(pct > 0, tab("red"), tab("blue"))) %>% 
  filter(month >= as.Date("2012-01-01"))

gg <- ggplot(toplot, aes(month, pct, color=colfill)) + 
  geom_point() + 
  scale_color_identity() +
  scale_y_continuous("Percent of published CVEs", label=label_percent(1), limits=c(0,1),
                     breaks=seq(0,1,0.2), 
                     expand=expansion(mult=c(0,0.01))) +
  scale_x_date("CVE Published (monthly)", date_breaks = "1 year", date_labels = "%Y") +
  labs(title = paste("Percent of Published CVEs with CVSS vectors + base score (all CNAs)")) +
  theme(plot.title = element_text(family=mainfont, face="italic", size=11))

gg
```

## Which CNAs are published more CVSS Vectors? 

This looks at the percent of CVEs with at least one valid CVSS record from each CNA and compares against the total number of CVEs from that CNA. 

```{r}
toplot <- metrics %>% 
  count(cna, has_cvss) %>% 
  spread(has_cvss, n, fill=0) %>% 
  mutate(total = no + yes, pct = yes / total) %>%
  mutate(colfill = ifelse(pct > 0, tab("red"), tab("blue")))

txtlab <- toplot %>% filter(yes > 0 & total > 100)

gg <- ggplot(toplot, aes(total, pct, label=cna)) +
  geom_text_repel(data=txtlab, color="gray40", segment.color="gray80",
                  size=6/.pt, family=mainfont, 
                  point.padding = 0, box.padding=0.05) +
  geom_point() +
  scale_y_continuous("Percent of CVEs with CVSS (by CNA)", label=label_percent(1), limits=c(0,1),
                     breaks=seq(0,1,0.2), 
                     expand=expansion(mult=c(0,0.01))) +
  labs(title = paste("Percent of Published CVEs by CNA, with CVSS vector + score")) +
  scale_x_log10("CVEs Published by CNA", label=mklab,
                breaks=10^seq(0,10), 
                minor_breaks=seq(1,9) * rep(10^seq(-5,10), each=9), 
                expand=expansion(mult=c(0.01, 0.01))) +
  theme(plot.title = element_text(family=mainfont, face="italic", size=11))

gg
```

## Completeness of CVSS data without MITRE as CNA 
```{r fig.alt="cvss_by_published_monthly_without_mitre"}

toplot <- metrics %>% 
  mutate(month = floor_date(published, unit="month")) %>% 
  filter(cna != "mitre") %>% 
  count(month, has_cvss) %>% 
  spread(has_cvss, n, fill=0) %>% 
  mutate(total = no + yes, pct = yes / total) %>%
  mutate(colfill = ifelse(pct > 0, tab("red"), tab("blue"))) %>% 
  filter(month >= as.Date("2012-01-01"))


gg <- ggplot(toplot, aes(month, pct, color=colfill)) + 
  geom_point() + 
  scale_color_identity() +
  scale_y_continuous("Percent of published CVEs", label=label_percent(1), limits=c(0,1),
                     breaks=seq(0,1,0.2), 
                     expand=expansion(mult=c(0,0.01))) +
  scale_x_date("CVE Published (monthly)", date_breaks = "1 year", date_labels = "%Y") +
  labs(title = paste("Percent of Published CVEs with CVSS vectors + base score (without MITRE)")) +
  theme(plot.title = element_text(family=mainfont, face="italic", size=11))

gg
```

## Last year of published CVEs: 

```{r}
toplot <- metrics %>% 
  filter(published >= today() %m-% years(1)) %>% 
  count(cna, has_cvss) %>% 
  spread(has_cvss, n, fill=0) %>% 
  mutate(total = no + yes, pct = yes / total) %>%
  mutate(colfill = ifelse(pct > 0, tab("red"), tab("blue")))

txtlab <- toplot %>% filter(yes > 0 & total > 100) %>% 
  mutate(pos = ifelse(pct > 0.5, "Over Half", "Under Half"))
good <- toplot %>% 
  filter(total > 100) %>% 
  filter(pct > 0.5) %>% 
  arrange(pct, total) %>%
  mutate(txt = paste0(yes, "/", total, " (", percent(pct, 0.1), ")")) %>% 
  mutate(cna = factor(cna, levels=cna, ordered=TRUE))

gg1 <- ggplot(good, aes(pct, cna, label=txt)) +
  geom_col() +
  geom_text(size=8/.pt, color="white", family=mainfont, hjust="right",
            nudge_x = -0.01) + 
  scale_x_continuous("Percent of CVEs with valid CVSS",
                     expand=c(0,0), label=label_percent(1), limits=c(0,1)) +
  labs(title = paste("CNAs with valid CVSS information")) +
  theme(plot.title = element_text(family=mainfont, face="italic", size=11),
        axis.title.y = element_blank(),
        axis.line.y = element_line(color="gray40", linewidth=0.5))

bad <- toplot %>% 
  # filter(total > 100) %>% 
  filter(pct <= 0.5) %>% 
  top_n(n = nrow(good), total) %>% 
  drop_na() %>% 
  arrange(pct, total) %>%
  mutate(txt = paste0(yes, "/", total, " (", percent(pct, 0.1), ")")) %>% 
  mutate(cna = factor(cna, levels=cna, ordered=TRUE))

gg2 <- ggplot(bad, aes(pct, cna, label=txt)) +
  geom_col() +
  geom_text(size=8/.pt, color="gray40", family=mainfont, hjust="left",
            nudge_x = 0.01) + 
  scale_x_continuous("Percent of CVEs with valid CVSS",
                     expand=c(0,0), label=label_percent(1), limits=c(0,1)) +
  labs(title = paste("CNAs missing CVSS information")) +
  theme(plot.title = element_text(family=mainfont, face="italic", size=11),
        axis.title.y = element_blank(),
        axis.line.y = element_line(color="gray40", linewidth=0.5))



gg1 + gg2 + plot_layout(nrow=1)

```


```{r eval=FALSE}
# validation
# missing 4.0 validation
version_regex <- c("2.0" = "^((AV:[NAL]|AC:[LMH]|Au:[MSN]|[CIA]:[NPC]|E:(U|POC|F|H|ND)|RL:(OF|TF|W|U|ND)|RC:(UC|UR|C|ND)|CDP:(N|L|LM|MH|H|ND)|TD:(N|L|M|H|ND)|[CIA]R:(L|M|H|ND))/)*(AV:[NAL]|AC:[LMH]|Au:[MSN]|[CIA]:[NPC]|E:(U|POC|F|H|ND)|RL:(OF|TF|W|U|ND)|RC:(UC|UR|C|ND)|CDP:(N|L|LM|MH|H|ND)|TD:(N|L|M|H|ND)|[CIA]R:(L|M|H|ND))$",
                   "3.0" = "^CVSS:3[.]0/((AV:[NALP]|AC:[LH]|PR:[UNLH]|UI:[NR]|S:[UC]|[CIA]:[NLH]|E:[XUPFH]|RL:[XOTWU]|RC:[XURC]|[CIA]R:[XLMH]|MAV:[XNALP]|MAC:[XLH]|MPR:[XUNLH]|MUI:[XNR]|MS:[XUC]|M[CIA]:[XNLH])/)*(AV:[NALP]|AC:[LH]|PR:[UNLH]|UI:[NR]|S:[UC]|[CIA]:[NLH]|E:[XUPFH]|RL:[XOTWU]|RC:[XURC]|[CIA]R:[XLMH]|MAV:[XNALP]|MAC:[XLH]|MPR:[XUNLH]|MUI:[XNR]|MS:[XUC]|M[CIA]:[XNLH])$",
                   "3.1" = "^CVSS:3[.]1/((AV:[NALP]|AC:[LH]|PR:[NLH]|UI:[NR]|S:[UC]|[CIA]:[NLH]|E:[XUPFH]|RL:[XOTWU]|RC:[XURC]|[CIA]R:[XLMH]|MAV:[XNALP]|MAC:[XLH]|MPR:[XNLH]|MUI:[XNR]|MS:[XUC]|M[CIA]:[XNLH])/)*(AV:[NALP]|AC:[LH]|PR:[NLH]|UI:[NR]|S:[UC]|[CIA]:[NLH]|E:[XUPFH]|RL:[XOTWU]|RC:[XURC]|[CIA]R:[XLMH]|MAV:[XNALP]|MAC:[XLH]|MPR:[XNLH]|MUI:[XNR]|MS:[XUC]|M[CIA]:[XNLH])$") 

tojoin <- version_regex %>% 
  enframe("key_version", "regex")

metrics <- raw_metrics %>% 
  select(cve, metrics_id, ends_with(c("vectorString", "version", "baseScore"))) %>% 
  distinct() %>% 
  pivot_longer(cols = -c(cve, metrics_id), names_to="key", values_to = "val", 
               values_transform = as.character, values_drop_na = TRUE) %>% 
  mutate(key = gsub("containers.cna.metrics.", "", key)) %>% 
  separate(key, into=c("key_version", "key_name"), sep="\\.") %>% 
  mutate(key_version = case_when(key_version == "cvssV3_1" ~ "3.1",
                                 key_version == "cvssV3_0" ~ "3.0",
                                 key_version == "cvssV2_0" ~ "2.0",
                                 TRUE ~ "unknown")) %>% 
  pivot_wider(id_cols = c(cve, metrics_id, key_version), names_from = key_name, values_from = val) %>% 
  full_join(cvepub, by="cve") %>% 
  full_join(cvecna, by="cve")

count_valid <- metrics %>% 
  filter(!is.na(vectorString)) %>% 
  left_join(tojoin, by="key_version") %>% 
  mutate(valid = str_detect(vectorString, regex)) %>% 
  count(cvss_vector_is_valid = valid)

count_valid
```
## Proportion of CVSS versions over time


```{r fig.alt = "cves_published_with_cvss_version"}

fancy_metrics <- raw_metrics %>% 
  select(cve, metrics_id, ends_with(c("vectorString", "version", "baseScore"))) %>% 
  distinct() %>% 
  pivot_longer(cols = -c(cve, metrics_id), names_to="key", values_to = "val", 
               values_transform = as.character, values_drop_na = TRUE) %>% 
  mutate(key = gsub("containers.cna.metrics.", "", key)) %>% 
  separate(key, into=c("key_version", "key_name"), sep="\\.") %>% 
  mutate(key_version = case_when(key_version == "cvssV3_1" ~ "3.1",
                                 key_version == "cvssV3_0" ~ "3.0",
                                 key_version == "cvssV2_0" ~ "2.0",
                                 TRUE ~ "unknown")) %>% 
  pivot_wider(id_cols = c(cve, metrics_id, key_version), names_from = key_name, values_from = val) %>% 
  full_join(cvepub, by="cve") %>% 
  full_join(cvecna, by="cve")

toplot <- fancy_metrics %>% 
  filter(!is.na(vectorString)) %>% 
  mutate(month = floor_date(published, unit="month")) %>%
  filter(month >= as.Date("2018-01-01")) %>% 
  distinct(cve, key_version, month) %>% 
  mutate(one = 1) %>% 
  spread(key_version, one, fill=0) %>% 
  summarize(.by=month, `2.0` = sum(`2.0`),
            `3.0` = sum(`3.0`),
            `3.1` = sum(`3.1`)) %>% 
  gather(cat, val, -month) %>% 
  left_join(cvemonth, by="month") %>%
  mutate(pct = val/n)

gg <- ggplot(toplot, aes(month,pct, group=cat, color=cat, label=cat)) +
  geom_line(alpha=1/4) +
  geom_point(alpha=1/2) +
  geomtextpath::geom_labelsmooth(size=8/.pt, family=mainfont, text_smoothing = 50, vjust=0.2, hjust=0.7,
                                 method = 'loess', formula = 'y ~ x') +
  scale_color_manual(values=c("2.0"=tab("red"), "3.0"=tab("seablue"), 
                              "3.1"=tab("blue"), "CVEs Published" = "gray60")) + 
  theme(legend.position = "none") +
  scale_y_continuous("Percent of Monthly Published CVEs", label=label_percent(1), 
                     expand=c(0,0), limits = c(0,1), breaks=seq(0,1,0.2))
gg
```


## Proportion of CVSS versions over time without MITRE as CNA 

```{r fig.alt = "cves_published_with_cvss_version_no_mitre"}
cvemonth_nomitre <- cves %>% 
  mutate(published = as.Date(str_extract(cveMetadata.datePublished, "\\d{4}-\\d{2}-\\d{2}"))) %>% 
  filter(cveMetadata.assignerShortName != "mitre") %>% 
  mutate(month = floor_date(published, unit="month")) %>% 
  count(month)

toplot <- fancy_metrics %>% 
  filter(cna != "mitre") %>% 
  filter(!is.na(vectorString)) %>% 
  mutate(month = floor_date(published, unit="month")) %>%
  filter(month >= as.Date("2017-01-01")) %>% 
  distinct(cve, key_version, month) %>% 
  mutate(one = 1) %>% 
  spread(key_version, one, fill=0) %>% 
  summarize(.by=month, `2.0` = sum(`2.0`),
            `3.0` = sum(`3.0`),
            `3.1` = sum(`3.1`)) %>% 
  gather(cat, val, -month) %>% 
  left_join(cvemonth_nomitre, by="month") %>%
  mutate(pct = val/n)

gg <- ggplot(toplot, aes(month,pct, group=cat, color=cat, label=cat)) +
  geom_line(alpha=1/10) +
  geom_point(alpha=1/2) +
  geomtextpath::geom_labelsmooth(size=8/.pt, family=mainfont, textsmoothing = 50, hjust=0.6) +
  scale_color_manual(values=c("2.0"=tab("red"), "3.0"=tab("seablue"), 
                              "3.1"=tab("blue"), "CVEs Published" = "gray60")) + 
  theme(legend.position = "none") +
  scale_y_continuous("Percent of Monthly Published CVEs", label=label_percent(1), 
                     expand=c(0,0), limits = c(0,1), breaks=seq(0,1,0.2))
gg
```


