---
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: false
    theme: paper
    code_folding: none
    toc_float:
        collapsed: false
        smooth_scroll: false
---

```{r setup, echo=FALSE}
suppressPackageStartupMessages({
  suppressWarnings({
    library(dplyr)
    library(tibble)
    library(tidyr)
    library(readr)
    library(stringr)
    library(here)
    library(scales)
    library(log4r)
    library(lubridate)
    library(purrr)
    library(arrow)
    library(tidyjson)
    library(ggplot2)
    library(ggrepel)
    library(gt)
    library(xml2)
    library(patchwork)
  })
})

mainfont <- "Source Sans Pro"
tab <- function (x = c("blue", "orange", "red", "seablue", "green", 
                       "olive", "purple", "pink", "brown", "gray")) {
  tableau <- c(blue = "#4E79A7", orange = "#F28E2B", red = "#E15759", seablue = "#76B7B2", 
               green = "#59A14F", olive = "#EDC948", purple = "#B07AA1", pink = "#FF9DA7", 
               brown = "#9C755F", gray = "#BAB0AC")
  as.vector(tableau[x])
}

theme_set(theme_minimal(base_family=mainfont) +
            theme(panel.grid = element_line(color="gray95"),
                  plot.caption = element_text(size=6, face = "italic", color="gray60"),
                  text = element_text(family=mainfont),
                  legend.title = element_blank(),
                  legend.position="bottom"))
update_geom_defaults('col', list(fill=tab('blue')))
update_geom_defaults('bar', list(fill=tab('blue')))
update_geom_defaults('text', list(family=mainfont, size=8/.pt))
update_geom_defaults('label', list(family=mainfont, size=8/.pt, label.size=NA))
# update_geom_defaults('point', list(shape=21, size=2, fill=tab('blue'), color='white'))
update_geom_defaults('point', list(size=0.5, color=tab('blue')))
update_geom_defaults('line', list(color=tab('blue'), size=0.85))

knitr::opts_chunk$set(
  cache = FALSE,
  message = FALSE,
  warning = FALSE, 
  dev = c("png", "cairo_pdf"),
  echo = FALSE,
  fig.retina = 2,
  fig.width = 7,
  fig.height = 3.5
)

```

```{r}

mklab <- function (x, bytes = FALSE) {
    brate <- tibble(cut = c(10^seq(3, 24, 3), Inf), suffix = c("0", 
        "K", "M", "B", "T", "q", "Q", "s", "S"))
    if (bytes) {
        brate <- tibble(cut = c(1024^seq(8), Inf), suffix = c("0", 
            "kB", "MB", "GB", "TB", "PB", "EB", "ZB", "YB"))
    }
    pickup <- tibble(x = x) %>% mutate(one = cut(x, breaks = c(-Inf, 
        brate$cut), labels = brate$suffix, right = FALSE)) %>% 
        mutate(two = x/sapply(.data$one, function(y) lag(brate$cut)[y])) %>% 
        mutate(three = ifelse(.data$two < 10, as.character(round(.data$two, 
            1)), as.character(round(.data$two, 0)))) %>% mutate(three = ifelse(is.na(.data$three), 
        as.character(round(.data$x, 0)), paste0(.data$three, 
            .data$one))) %>% mutate(four = ifelse(x < 1, x, .data$three))
    pickup$four
}

```

```{r}
cves <- read_parquet(here("cache/cveparse/cves.parquet")) %>% 
  select(-contains(".x_"))
cvepub <- cves %>% 
  select(cve, published = cveMetadata.datePublished) %>% drop_na() %>% 
  mutate(published = as.Date(str_extract(published, "\\d{4}-\\d{2}-\\d{2}")))
cvemonth <- cvepub %>% 
  mutate(month = floor_date(published, unit="month")) %>% 
  count(month)
cvecna <- cves %>% 
  select(cve, cna = cveMetadata.assignerShortName) %>% 
  drop_na()
cnacount <- cves %>% 
  select(cve, cna = cveMetadata.assignerShortName) %>% 
  drop_na() %>%
  count(cna, name="n")
totalcve <- cves %>% distinct(cve) %>% nrow()
totalcna <- cves %>% distinct(cveMetadata.assignerShortName) %>% nrow()

```

# Problem types

```{r}
raw_ptype <- read_parquet(here("cache", "cveparse", "containers.cna.problemTypes.parquet")) %>% 
  left_join(read_parquet(here("cache", "cveparse", "containers.cna.problemTypes.descriptions.parquet")), 
            by=c("cve", "problemTypes_id")) %>% 
  mutate(across(where(is.character), function(x) gsub("^n/a$", NA_character_, x)))

cleaned <- raw_ptype %>% 
  rename_with(~ gsub("containers.cna.problemTypes.descriptions.", "", .x, fixed = TRUE))

fortable <- cleaned %>% 
  gather(cat, val, -cve) %>% 
  filter(!cat %in% c("problemTypes_id", "descriptions_id")) %>% 
  drop_na() %>% 
  mutate(strlen = nchar(cat)) %>% 
  left_join(cvecna, by="cve") %>% 
  mutate(cat = paste0("problemTypes.", cat)) %>% 
  arrange(strlen, cat) %>% 
  summarize(.by=cat, cves = n_distinct(cve), cnas = n_distinct(cna)) 

knitr::kable(fortable, format = "html", 
             caption = paste(cves %>% distinct(cve) %>% nrow() %>% comma, 
                     "CVEs from", comma(totalcna), "CNAs as of ", 
                     as.Date(max(cves$cveMetadata.datePublished, na.rm = TRUE))))

```

## "Type" in problemTypes

```{r}
# type is interesting.
toplot <- cleaned %>% 
  left_join(cvecna, by="cve") %>% 
  summarize(.by=type, cves = n_distinct(cve), cnas = n_distinct(cna)) %>% 
  arrange(-cves) %>% 
  replace_na(list(type = "<undefined>")) %>% 
  mutate(type = factor(type, levels=type, ordered=TRUE))

only_one <- toplot %>% 
  filter(cnas <= 2) %>% 
  pull(type) %>% 
  as.character()
toplot2 <- cleaned %>% 
  left_join(cvecna, by="cve") %>% 
  filter(type %in% only_one) %>% 
  summarize(.by=type, cna = paste(unique(cna), collapse=", ")) %>% 
  replace_na(list(type = "<undefined>")) %>% 
  mutate(type = factor(type, levels=levels(toplot$type), ordered=TRUE)) %>% 
  right_join(toplot, by="type") %>% 
  arrange(type) %>% 
  replace_na(list(cna = "<too many>")) %>% 
  select(type, cves, cnas, cna)

knitr::kable(toplot2, format = "html", 
             caption = paste("problemType.type values and counts"))
```

## Looking for CWE through regex

I used a regex "cwe\[-\_.\]\\d+" to detect a CWE ID in any of the `description` or the two `cweId` and `cweid` fields. The following table is a count of the CVEs with one or more CWEs found in those fields where the type is specified as shown.

```{r}
toplot <- cleaned %>% 
  select(cve, description, lang, type, cweId) %>% 
  mutate(in_desc = str_detect(tolower(description), "cwe[-_.]\\d+"),
         in_cweId = str_detect(tolower(cweId), "cwe[-_.]\\d+")) %>% 
  replace_na(list(in_desc=FALSE, in_cweId=FALSE)) %>% 
  mutate(anycwe = in_desc | in_cweId) %>% 
  # filter(anycwe) %>%
  left_join(cvecna, by="cve") %>% 
  summarize(.by=c(type,anycwe), cves = n_distinct(cve)) %>% 
  mutate(label = ifelse(anycwe, "CWE Found", "No CWE")) %>% 
  select(type, label, cves) %>% 
  spread(label, cves, fill=0)

knitr::kable(toplot, format = "html", 
             caption = paste("The elusive CWE ID"))

```

Because it's a simple regex we can discover the CWE ID in other places, such as the description. This is limited to the records where a valid CWE field was found.

```{r}
toplot <- cleaned %>% 
  select(cve, description, lang, type, cweId) %>% 
  mutate(in_desc = str_detect(tolower(description), "cwe[-_.]\\d+"),
         in_cweId = str_detect(tolower(cweId), "cwe[-_.]\\d+")) %>% 
  replace_na(list(in_desc=FALSE, in_cweId=FALSE)) %>% 
  mutate(anycwe = in_desc | in_cweId) %>% 
  filter(anycwe) %>%
  left_join(cvecna, by="cve") %>% 
  summarize(.by=c(in_desc, in_cweId), cves = n_distinct(cve), cnas=n_distinct(cna)) %>% 
  arrange(-cves, -cnas) %>% 
  rename(`CWE in Description`=in_desc,
         `CWE in cweId`=in_cweId)

knitr::kable(toplot, format = "html", 
             caption = paste("Where are CWEs being found?"))

```

## Record Completeness

Should there always be a "lang" and "type" and "cweId" or what's the best combination?

```{r}
toplot <- cleaned %>% 
  gather(cat, val, -cve) %>% 
  filter(!cat %in% c("problemTypes_id", "descriptions_id")) %>% 
  drop_na() %>% 
  mutate(strlen = nchar(cat)) %>% 
  # left_join(cvecna, by="cve") %>% 
  mutate(cat = paste0("problemTypes.", cat)) %>% 
  arrange(strlen, cat) %>% 
  mutate(cat = factor(cat, levels=unique(cat), ordered=TRUE)) %>% 
  # complete(cve,cat) %>% 
  mutate(value = !is.na(val)) %>%
  summarize(.by=c(cve, cat), value = any(value)) %>% 
  # complete(cve,cat) %>% 
  # replace_na(list(value=FALSE)) %>% 
  # count(cat, value)
  pivot_wider(id_cols = cve, names_from = cat, values_from = value, values_fill = FALSE) %>% 
  left_join(cvepub, by="cve") %>% 
  replace_na(list(problemTypes.lang = FALSE, problemTypes.type = FALSE, problemTypes.cweId = FALSE, 
        problemTypes.description = FALSE)) %>% 
  count(lang=problemTypes.lang, type=problemTypes.type, 
        cweId=problemTypes.cweId, 
        description=problemTypes.description, name="count") %>% 
  arrange(count) %>% 
  mutate(yval = row_number()) %>% 
  gather(cat, val, -yval) %>% 
  mutate(cat = factor(cat, levels=c("count", "lang", "type", "cweId", "description"), ordered=TRUE)) %>% 
  mutate(filler = ifelse(val == 1 & cat != "count", tab("blue"), "white"),
         val = case_when(val == 1 & cat != "count" ~ "X",
                         val == 0 & cat != "count" ~ "",
                         TRUE ~ comma(as.numeric(val))))

gg <- ggplot(toplot, aes(cat, yval, label=val, fill=filler)) +
  geom_tile(color="gray95", size=1) +
  geom_text(size=8/.pt, family=mainfont) +
  scale_fill_identity() +
  scale_x_discrete(expand=c(0,0), position="top") +
  scale_y_continuous(expand=c(0,0)) +
  theme(axis.text.y = element_blank(),
        axis.title = element_blank())

gg


```

```{r eval=FALSE}
toplot <- cleaned %>% 
  count(description, lang, type, cweId, name="instances", sort=TRUE) %>% 
  mutate(in_desc = str_detect(tolower(description), "cwe[-_.]\\d+"),
         in_cweId = str_detect(tolower(cweId), "cwe[-_.]\\d+")) %>% 
  replace_na(list(in_desc=FALSE, in_cweId=FALSE)) %>% 
  mutate(in_cweId = in_cweId) %>% 
  count(in_desc, in_cweId, wt=instances, sort=TRUE)
```

# CWE ID

```{r}
zipfile <- tempfile(fileext = ".xml.zip")
download.file("https://cwe.mitre.org/data/xml/cwec_latest.xml.zip", zipfile)
outdir <- tempdir()
rez <- unzip(zipfile, exdir = outdir)
rawcwe <- read_xml(rez)
allcwe <- lapply(xml_children(rawcwe), function(cwecat) {
  # cat("Name:", xml_name(cwecat), "\n")
  allsubcwe <- map_dfr(xml_children(cwecat), function(subcwe) {
    basedata <- tibble(toplevel = xml_name(cwecat))

    basedata <- basedata %>% 
      bind_cols(xml_attrs(subcwe) %>% enframe() %>% 
                  spread(name, value))
    # cat("------------", basedata$ID, "\n")
    # print(glimpse(basedata))
    for (weak in xml_children(subcwe)) {
      if(xml_name(weak) == "Related_Weaknesses") {
        related <- map_dfr(xml_children(weak), function(subweak) {
          xml_attrs(subweak) %>% enframe() %>% 
            spread(name, value) 
        }) %>% janitor::clean_names()
        basedata <- basedata %>% bind_cols(tibble(Releated_Weaknesses = related))
      } else if (xml_name(weak) == "Relationships") {
        relations <- map_dfr(xml_children(weak), function(subweak) {
          xml_attrs(subweak) %>% enframe() %>% 
            spread(name, value) %>% 
            mutate(xmlname = xml_name(subweak))
        }) %>% janitor::clean_names()
        basedata <- basedata %>% bind_cols(tibble(Relationships = relations))
      } else {
        if(length(xml_attrs(weak)) > 0) {
          basedata <- basedata %>% 
            bind_cols(xml_attrs(weak) %>% enframe() %>% 
                        mutate(name = paste0(xml_name(weak), ".", name)) %>% 
                        spread(name, value))
        }
        if(nchar(xml_text(weak)) > 0) {
          textval <- xml_text(weak)
          names(textval) <- xml_name(weak)
          basedata <- basedata %>% 
            bind_cols(textval %>% enframe() %>% spread(name, value))
        }
      }
    }
    basedata
  }) %>% janitor::clean_names()
  allsubcwe
}) 
weakness <- allcwe[[1]] %>% select(-any_of("related_weaknesses"))
related_weakness <- allcwe[[1]] %>% 
  select(id, releated_weaknesses) %>% 
  unnest(releated_weaknesses) %>% 
  filter(!is.na(cwe_id))
category <- allcwe[[2]]  %>% select(-any_of("relationships"))
category_relation <- allcwe[[2]] %>% 
  select(id, relationships) %>% 
  unnest(relationships) %>%
  left_join(category %>% select(toplevel, id, name, status), by="id") %>% 
  distinct()

views <- allcwe[[3]]
external_ref <- allcwe[[4]]
cwe_view <- allcwe[[3]]

cwe_map <- weakness %>% select(toplevel, id, name, status, abstraction) %>% 
  bind_rows(category %>% select(toplevel, id, name, status)) %>% 
  bind_rows(views %>% select(toplevel, id, name, status)) %>% 
  distinct() %>% 
  mutate(toplevel = case_when(toplevel == "Weaknesses" ~ "Weakness",
                              toplevel == "Categories" ~ "Category",
                              toplevel == "Views" ~ "View")) %>% 
  mutate(txt = ifelse(!is.na(abstraction), abstraction, toplevel))

# cwe_count <- cwe %>% 
#   mutate(id = str_extract(cwe, "\\d+")) %>% 
#   left_join(cwe_map, by="id") %>% 
#   count(cve, toplevel) %>% 
#   mutate(txt = paste0(toplevel, "(", n, ")")) %>% 
#   summarize(.by=cve, x = paste(sort(txt), collapse=",")) %>% 
#   count(x, name="cves", sort=TRUE) %>% 
#   mutate(pct = percent(cves/sum(cves), 0.1))
# cwe_count
```

## Number off CWEs in CVE records

This section is aggregating any identification of CWE ID in any of the three common fields.

Note that seven CVEs had five (5) unique CWEs in a single CVE record.

```{r fig.width=7, fig.height=2}
toplot <- cleaned %>% 
  select(cve, description, lang, type, cweId) %>% 
  mutate(in_desc = str_extract(tolower(description), "cwe[-_.]\\d+"),
         in_cweId = str_extract(tolower(cweId), "cwe[-_.]\\d+")) %>% 
  filter(!(is.na(in_desc) & is.na(in_cweId))) %>% 
  mutate(cwe = case_when(!is.na(in_cweId) ~ toupper(in_cweId),
                         !is.na(in_desc) ~ toupper(in_desc))) %>% 
  filter(!is.na(cwe)) %>% 
  distinct(cve, cwe) %>% 
  count(cve, name="cwes") %>% 
  count(cwes, name="cves") %>% 
  mutate(cwes = factor(cwes, levels=rev(sort(unique(cwes))), ordered=TRUE)) %>% 
  mutate(txt = paste(comma(cves), "CVEs"))


gg <- ggplot(toplot, aes(cves, cwes, label=txt)) +
  geom_vline(xintercept=0, color="gray90", size=0.25) +
  geom_col(fill=tab("blue")) +
  ggfittext::geom_bar_text(size=8) +
  scale_y_discrete("Number of unique\nCWEs in CVE record", expand=c(0,0)) +
  scale_x_continuous("Count of CVEs", expand=c(0,0)) +
  theme(axis.text.x = element_blank(),
        panel.grid = element_blank())
gg
```

CWE's have different level of Abstractions as well as views and categories. CWE recommends only assigning "Base" CWEs to vulnerabilities.

```{r}
toplot <- cleaned %>% 
  select(cve, description, lang, type, cweId) %>% 
  mutate(in_desc = str_extract(tolower(description), "cwe[-_.]\\d+"),
         in_cweId = str_extract(tolower(cweId), "cwe[-_.]\\d+")) %>% 
  filter(!(is.na(in_desc) & is.na(in_cweId))) %>% 
  mutate(cwe = case_when(!is.na(in_cweId) ~ toupper(in_cweId),
                         !is.na(in_desc) ~ toupper(in_desc))) %>% 
  filter(!is.na(cwe)) %>% 
  distinct(cve, cwe) %>% 
  mutate(id = gsub("CWE-", "", cwe)) %>% 
  # left_join(cvecna, by="cve") %>% 
  left_join(cwe_map, by=c("id")) %>% 
  mutate(year = as.numeric(substr(cve, 5, 8))) %>% 
  filter(!is.na(txt)) %>% 
  count(year, txt) %>% 
  filter(year >= 2017) %>% 
  mutate(.by=year, pct = n/sum(n))

toplot2 <- toplot %>% count(txt, wt=n) %>% 
  mutate(pct = n/sum(n)) %>% 
  arrange(pct) %>% 
  mutate(txt = factor(txt, levels=txt, ordered=TRUE),
         lab = percent(pct, 0.1))

toplot3 <- toplot %>% 
  mutate(txt = factor(txt, levels=levels(toplot2$txt), ordered=TRUE))

colmap <- rev(ggthemes::tableau_color_pal()(length(levels(toplot2$txt))))
names(colmap) <- levels(toplot2$txt)

gg1 <- ggplot(toplot2, aes(pct, txt, fill=txt, label=lab)) +
  geom_col() +
  ggfittext::geom_bar_text(family=mainfont, size=9) +
  scale_fill_manual(values=colmap) +
  scale_y_discrete("CWE Abstraction", expand=c(0,0)) +
  scale_x_continuous("Percent of CWE values", expand=c(0,0), label=label_percent(1))
  #theme(axis.title.y = element_blank())
  
  
gg2 <- ggplot(toplot3, aes(year, pct, group=txt, fill=txt)) +
  geom_col(position="stack") +
  scale_fill_manual(values=colmap) +
  scale_x_continuous("CVE Year", expand=c(0,0), breaks=seq(2017,2030)) +
  scale_y_continuous("Percent of CWE values", expand=c(0,0), label=label_percent(1))

gg1 + gg2 + plot_layout(nrow=1, guides = "collect") &
    scale_fill_manual(values=colmap)

```

## top CWEs

```{r fig.height=9}
toplot <- cleaned %>% 
  select(cve, description, lang, type, cweId) %>% 
  mutate(in_desc = str_extract(tolower(description), "cwe[-_.]\\d+"),
         in_cweId = str_extract(tolower(cweId), "cwe[-_.]\\d+")) %>% 
  filter(!(is.na(in_desc) & is.na(in_cweId))) %>% 
  mutate(cwe = case_when(!is.na(in_cweId) ~ toupper(in_cweId),
                         !is.na(in_desc) ~ toupper(in_desc))) %>% 
  filter(!is.na(cwe)) %>% 
  distinct(cve, cwe) %>% 
  mutate(id = gsub("CWE-", "", cwe)) %>% 
  # left_join(cvecna, by="cve") %>% 
  left_join(cwe_map, by=c("id")) %>% 
  filter(!is.na(name)) %>% 
  mutate(txt = paste0(cwe, ": ", str_trunc(name, width=40, side = "right"))) %>% 
  count(txt) %>% 
  arrange(n) %>% 
  mutate(txt = factor(txt, levels=txt, ordered=TRUE)) %>% 
  mutate(pct = n/sum(n),
         lab = paste0(n, " (", percent(pct, 0.1), ")")) %>% 
  top_n(50, wt=n)

gg <- ggplot(toplot, aes(pct, txt, label=lab)) +
  geom_col(color=NA) +
  ggfittext::geom_bar_text(size=7, min.size = 5, family=mainfont, outside=TRUE) +
  scale_x_continuous("Percent of CWEs", expand=c(0,0)) +
  theme(axis.title.y = element_blank(),
        axis.line.y = element_line(color="gray60", size=0.5),
        panel.grid = element_blank())
gg
```

We can get a better idea of what these are by using the "Comprehensive Categorization" in category 1400. It maps every CWE up to a one of twenty-two high-level categories.

```{r fig.height=6}
toplot <- cleaned %>% 
  select(cve, description, lang, type, cweId) %>% 
  mutate(in_desc = str_extract(tolower(description), "cwe[-_.]\\d+"),
         in_cweId = str_extract(tolower(cweId), "cwe[-_.]\\d+")) %>% 
  filter(!(is.na(in_desc) & is.na(in_cweId))) %>% 
  mutate(cwe = case_when(!is.na(in_cweId) ~ toupper(in_cweId),
                         !is.na(in_desc) ~ toupper(in_desc))) %>% 
  filter(!is.na(cwe)) %>% 
  distinct(cve, cwe) %>% 
  mutate(id = gsub("CWE-", "", cwe))
  # left_join(cvecna, by="cve") %>% 

toplot <- category_relation %>% 
  filter(view_id == "1400") %>% 
  rename(cat_id = id, id = cwe_id) %>% 
  select(cat_id, id, name) %>% 
  mutate(name = gsub("Comprehensive Categorization: ", "", name)) %>% 
  right_join(toplot, by="id") %>% 
  filter(!is.na(name)) %>% 
  # summarize(.by=cve, name = paste(sort(unique(name)), collapse=", ")) %>% 
  count(name, sort=TRUE) %>% 
  arrange(n) %>% 
  mutate(name = factor(name, levels=name, ordered=TRUE)) %>% 
  mutate(pct = n/sum(n),
         lab = paste0(n, " (", percent(pct, 0.1), ")")) %>% 
  top_n(50, wt=n)

gg <- ggplot(toplot, aes(pct, name, label=lab)) +
  geom_col(color=NA) +
  ggfittext::geom_bar_text(size=7, min.size = 4, family=mainfont, outside=TRUE) +
  scale_x_continuous("Percent of CWEs", expand=c(0,0)) +
  theme(axis.title.y = element_blank(),
        axis.line.y = element_line(color="gray60", size=0.5),
        panel.grid = element_blank())
gg


```

```{r}
toplot <- cleaned %>% 
  select(cve, description, lang, type, cweId) %>% 
  mutate(in_desc = str_extract(tolower(description), "cwe[-_.]\\d+"),
         in_cweId = str_extract(tolower(cweId), "cwe[-_.]\\d+")) %>% 
  filter(!(is.na(in_desc) & is.na(in_cweId))) %>% 
  mutate(cwe = case_when(!is.na(in_cweId) ~ toupper(in_cweId),
                         !is.na(in_desc) ~ toupper(in_desc))) %>% 
  filter(!is.na(cwe)) %>% 
  distinct(cve, cwe) %>% 
  mutate(id = gsub("CWE-", "", cwe))

nvdview <- read_csv(here("cache", "1003.csv"), show_col_types = FALSE) %>% 
  select(id = `CWE-ID`, 
         name=Name, 
         abstraction=`Weakness Abstraction`, 
         related=`Related Weaknesses`) %>% 
  mutate(id = as.character(id))

toplot1 <- toplot %>% 
  left_join(nvdview, by=c("id"))

in_cves <- nrow(toplot1)
in_1003 <- toplot1 %>% filter(is.na(name)) %>% nrow()

pct_in_nvd <- percent(in_1003 / in_cves, 0.1)
```

Out of the `r comma(in_cves)` CWEs used in the in the CVEs, `r comma(in_1003)` are identified to be part of the CWE View 1003 (the "NVD Slice").  That is `r pct_in_nvd` of the CWEs found in CVEs. 


```{r}
toplot <- cleaned %>% 
  select(cve, description, lang, type, cweId) %>% 
  mutate(in_desc = str_extract(tolower(description), "cwe[-_.]\\d+"),
         in_cweId = str_extract(tolower(cweId), "cwe[-_.]\\d+")) %>% 
  filter(!(is.na(in_desc) & is.na(in_cweId))) %>% 
  mutate(cwe = case_when(!is.na(in_cweId) ~ toupper(in_cweId),
                         !is.na(in_desc) ~ toupper(in_desc))) %>% 
  filter(!is.na(cwe)) %>% 
  distinct(cve, cwe) %>% 
  mutate(id = gsub("CWE-", "", cwe))

fixdate <- read_parquet(here("../fixed_cve_published_dates_2024-08-27.parquet")) %>% 
  select(cve, realdate=published)
has_cwe <- cvepub %>% 
  left_join(fixdate, by="cve") %>% 
  mutate(published = if_else(!is.na(realdate), realdate, published)) %>% 
  full_join(cvecna, by="cve") %>% 
  mutate(cwe = cve %in% toplot$cve) %>% 
  select(cve, published, cna, cwe)
  


alldates <- seq(as.Date("2017-01-01"), today(), by="2 days")

allplot <- map_dfr(alldates, function(curdate) {
  one <- has_cwe %>% 
    filter(published > (curdate - days(90)),
           published <= curdate)
  one %>% 
    summarize(x = sum(cwe), n=n(), pct = x/n, date = curdate, src="All CNAs") %>% 
    bind_rows(one %>% 
                filter(cna != "mitre") %>% 
                summarize(x = sum(cwe), n=n(), pct = x/n, date = curdate, src="Without MITRE"))
})

gg1 <- ggplot(allplot, aes(date, pct, group=src, color=src, label=src)) + 
  geom_line(alpha=1/4) +
  geomtextpath::geom_textpath(text_smoothing=50, vjust=0) +
  scale_y_continuous("Percent of Published CVEs with any CWE defined", label=label_percent(1),
                     limits=c(0,1)) +
  scale_color_manual(values=c("All CNAs"=tab("seablue"), "Without MITRE"=tab("blue"))) +
  scale_x_date("CVE published (showing end date of 90-day sliding window)",
               date_breaks = "1 year", date_labels = "%Y") +
  theme(legend.position="none")
gg1
```


